---
title: "PBMC multi-modal reference mapping"
always_allow_html: yes
date: "Compiled `r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    theme: united
    code_folding: hide
    collapse: yes
    css: max.css
---

[BACK TO INDEX](./index.html)

```{r, message=F, warning=F}
set.seed(123)
knitr::opts_knit$set(progress=F, verbose=F, warning=F, message=F)

library(Seurat)
library(tidyverse)
library(DT)
library(data.table)
library(ggplot2)
library(SeuratDisk)
library(ggpubr)
library(rstatix)
library(celda)
library(viridis)
library(ComplexHeatmap)
library(circlize)
library(ggvenn)
# library(RColorBrewer)

library(future)
plan("multicore", workers=18)

dontshow <- sapply(c("dplyr","Seurat", "HGNChelper", "openxlsx"), library, character.only=T, quietly=T)
source("gene_sets_prepare.R"); source("sctype_score_.R"); source("auto_detect_tissue_type.R")
source("aux.R")
```

# Blood (PBMC)

## Meta data information
```{r}
meta = read.csv('meta2.csv', sep='\t', header=T, stringsAsFactors=T)
meta$Health = factor(meta$Health, levels=c('HC', 'PV', 'PSORI', 'BP'), labels=c('HC', 'PV', 'PSORI', 'BP'))
meta = subset(meta, Tissue=='pbmc')
datatable(meta)
meta$PVvsHC = factor(ifelse(meta$Health %in% c('PV', 'HC'), as.character(meta$Health), NA), levels=c('PV', 'HC'))
meta$BPvsHC = factor(ifelse(meta$Health %in% c('BP', 'HC'), as.character(meta$Health), NA), levels=c('BP', 'HC'))
meta$PSORIvsHC = factor(ifelse(meta$Health %in% c('PSORI', 'HC'), as.character(meta$Health), NA), levels=c('PSORI', 'HC'))
```


```{r warning=F}
counts = fread("data/Combined_S3-forced-pbmc_DBEC_MolsPerCell.csv", sep=",", header=TRUE) %>% data.frame(row.names=1)
ex3 = readRDS('archive/Ex3-wta_Seurat.rds')
ggvenn(list(s3=rownames(counts), ex3=rownames(ex3@meta.data)), auto_scale=T)
genes = !grepl("pAbO", names(counts)) # select only gene columns
num_rna = sum(genes==T)
num_ab = sum(genes==F)
counts = counts %>% dplyr::rename(VSIR=C10orf54, SCYL1=CCL13, IGHA1=IGHA1_secreted, IGHD=IGHD_membrane,
                  IGHA1=IGHA1_secreted, IGHG1=IGHG1_secreted, IGHG2=IGHG2_secreted,
                  IGHG3=IGHG3_secreted, IGHG4=IGHG4_secreted, IGHM=IGHM_secreted,
                  TRGC1=TARP_refseq, FYB1=FYB, `HLA-A`=HLA.A, `HLA-DMA`=HLA.DMA,
                  `HLA-DPA1`=HLA.DPA1, `HLA-DQB1`=HLA.DQB1, `HLA-DRA`=HLA.DRA,
                  `HLA-DR`=HLA.DR.CD74.AHS0035.pAbO, `CD56-1`=CD56.NCAM16.2.NCAM1.AHS0019.pAbO,
                  CD185=CXCR5.CXCR5.AHS0039.pAbO, CD186=CXCR6.CXCR6.AHS0148.pAbO,
                  CD357=GITR.TNFRSF18.AHS0104.pAbO, CD366=Tim3.HAVCR2.AHS0016.pAbO,
                  `CD4-1.pAb0`=CD4.SK3.CD4.AHS0032.pAbO, `CD3-1.pAb0`=CD3.UCHT1.CD3E.AHS0231.pAbO)
names(counts)[!genes] = gsub('\\..*$', '', names(counts)[!genes]) # rename antibody columns

counts = counts[order(row.names(counts)),]
pbmc <- CreateSeuratObject(counts=t(counts[,genes]), assay="RNA", project='pbmc')
pbmc[['AB']] = CreateAssayObject(counts=t(counts[,!genes]))
rm(counts)

# original sample calls for each cell
cell_assign <- fread("data/S3-forced-pbmc_Sample_Tag_Calls.csv", sep=",", header=TRUE) %>% data.frame(row.names=1)
cell_assign = cell_assign[order(row.names(cell_assign)),]

# Reads of each sample per cell
cell_sample_reads <- fread("data/S3-forced-pbmc_Sample_Tag_ReadsPerCell.csv", sep=",", header=TRUE) %>% data.frame(row.names=1)
cell_sample_reads = cell_sample_reads[order(row.names(cell_sample_reads)),]
names(cell_sample_reads) = gsub('\\.stAbO', '', names(cell_sample_reads))

stopifnot(identical(row.names(cell_assign), row.names(cell_sample_reads)))

cell_sample_recalls = data.frame(col=apply(cell_sample_reads, 1, function(r) { which.max(r) }), og_tag=cell_assign$Sample_Tag, og_sample=cell_assign$Sample_Name)
# cell_sample_recalls$Sample_Tag = cell_assign$Sample_Tag
cell_sample_recalls$Sample_Tag = names(cell_sample_reads)[cell_sample_recalls$col]
cell_sample_recalls$ratio = apply((cell_sample_reads+1)/colSums(cell_sample_reads), 1, function(r) max(r)/sum(r)) # normalized by total of each sample tag
cell_sample_recalls$Max_Tag_Reads = apply(cell_sample_reads, 1, max)

cell_sample_recalls = merge(cell_sample_recalls, meta, by='Sample_Tag', all.x=T)
row.names(cell_sample_recalls) = row.names(cell_assign)
cell_sample_recalls$multiplet = factor(ifelse(cell_sample_recalls$og_sample=='Multiplet', 'multiplet', ifelse(cell_sample_recalls$og_sample=='Undetermined', 'undetermined', 'unique')))
pbmc <- AddMetaData(pbmc, metadata=cell_sample_recalls)
# pbmc = subset(pbmc, !Sample_Name %in% c('Multiplet', 'Undetermined'))
rm(cell_sample_recalls); rm(cell_sample_reads); rm(cell_assign)


counts <- LayerData(pbmc, assay = "RNA", layer = "counts")
counts <- Matrix::t(Matrix::t(counts)/Matrix::colSums(counts)) * 100
most_expressed <- order(apply(counts, 1, median), decreasing = T)[10:1]
boxplot(as.matrix(t(counts[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell",
    col = rainbow(length(most_expressed)), horizontal=T, main='top transcripts')

counts <- LayerData(pbmc, assay = "AB", layer = "counts")
counts <- Matrix::t(Matrix::t(counts)/Matrix::colSums(counts)) * 100
most_expressed <- order(apply(counts, 1, median), decreasing = T)[10:1]
boxplot(as.matrix(t(counts[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell",
    col = rainbow(length(most_expressed)), horizontal=T, main='top surface markers')

pbmc@meta.data$Health = factor(pbmc@meta.data$Health, levels=c('HC', 'PV', 'PSORI', 'BP'), labels=c('HC', 'PV', 'PSORI', 'BP'))
```


## Data quality
Ideally, cells should have a large number of reads overall (x-axis) and fairly broad coverage of a large number of genes (y-axis).
Cells with few detected genes (bottom) make normalization and imputation hard to impossible.
We can use minimal thresholds (red dashed lines) for number of genes and number of overall reads to exclude low quality cells.
```{r, fig.width=9, fig.height=9, warning=F}
feature_min = 10
read_min = 10
ratio_min = 0.5
ggplot(pbmc@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=multiplet, alpha=ratio)) + 
  geom_point(shape=16, size=1) + scale_color_manual(values=list(undetermined='orange', multiplet='red', unique='navy')) +
  scale_alpha(range=c(0.8, 0.1)) + scale_x_log10()  + theme_classic() +
  geom_hline(yintercept=feature_min, color='red', lty='dashed') +
  geom_vline(xintercept=read_min, color='red', lty='dashed') +
  facet_wrap(~Sample_Name, ncol=3)
```

Notably, multiplets are scattered throughout the distribution.
I assume uniquely identified cells also just contain reads from one sample tags, whereas remaining multiplets count reads from two or more tags that we cannot disentangle at this point.

```{r, fig.width=11, fig.height=5}
ggplot(pbmc@meta.data, aes(color=Sample_Name, x=nCount_RNA, fill=Sample_Name)) +
   geom_density(alpha=0.2) + scale_x_log10() + theme_classic() +
   geom_vline(xintercept=read_min, color='red', lty='dashed') + facet_wrap(~multiplet, ncol=1) +
   labs(title='UMI counts per cell', y='cell density') + scale_y_continuous(expand=c(0,0)) +

   ggplot(pbmc@meta.data, aes(color=Sample_Name, x=nFeature_RNA, fill=Sample_Name)) +
   geom_density(alpha=0.2) + scale_x_log10() + theme_classic() +
   labs(y="Cell density", title='Genes per cell') + facet_wrap(~multiplet, ncol=1) +
   geom_vline(xintercept=feature_min, color='red', lty='dashed') + scale_y_continuous(expand=c(0,0)) + theme(legend.position='none')
```


### Multiplet analysis
Based on the most prominent sample tag per cell, multiplets could be "rescued".
When plotting the ratio of the most prominent sample tag among all reads, we see that most multiplets are dominated by one tag.
```{r, fig.width=11, fig.height=5, warning=F}
ggplot(subset(pbmc@meta.data, multiplet!='undetermined'), aes(x=nCount_RNA, y=ratio, color=factor(Replicate)), alpha=0.1) +
  geom_point(shape=16, size=1) +
  scale_x_log10() + theme_classic() +
  geom_vline(xintercept=read_min, color='red', lty='dashed') +
   geom_hline(yintercept=ratio_min, color='red', lty='dashed') +
  facet_grid(vars(multiplet), vars(Health))
```

Including cells with at least the absolute majority of one sample (50% of tags) seems feasible but poses the risk of introducing "heterogenous" (more variable) cells to the analysis and muddling the clustering.
Note that some of the uniquely identified cells also contain more than 50% other sample tags.

⚠ Due to sufficiently high cell numbers in the PBMC samples overall, we decided to not recover multiplet in this dataset!

```{r, fig.width=11, fig.height=5}
p1 = ggplot(pbmc@meta.data, aes(x=factor(Sample_Name), fill=Sample_Name, alpha=multiplet)) + scale_alpha_ordinal(range=c(0.2, 1)) +
   geom_bar(show.legend=F) + coord_flip() + labs(y='# cells', x='', title='Cells per sample (+ multiplet rescue)') + theme_classic() + scale_y_continuous(expand=c(0.002, 0.002))

p2 = ggplot(subset(pbmc@meta.data, multiplet!='unique'), aes(x=og_sample, alpha=multiplet)) + scale_alpha_ordinal(range=c(0.8, 0.9))+
   geom_bar(aes(fill=Sample_Name), show.legend=F) + labs(y='# cells', x='', title='Cells per sample') + theme_classic() + scale_y_continuous(expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))

ggarrange(p1, p2, widths=c(3,1))
```

## Multimodal UMAP
```{r, echo=F, warning=F, message=F}
# pbmc@meta.data$Sample_Name = pbmc@meta.data$og_Sample
pbmc <- subset(pbmc, subset=nFeature_RNA >= feature_min & nCount_RNA >= read_min & multiplet=='unique')# & ABcontamination < 0.75) #ratio > 0.5  & pbmc$multiplet=='unique'

DefaultAssay(pbmc) <- "RNA"
pbmc = pbmc %>%
   NormalizeData(verbose=F) %>%
   FindVariableFeatures(verbose=F) %>%
   ScaleData() %>%
   RunPCA(reduction.name='rna.pca', reduction.key="rnaPC_", npcs=35, approx=F, verbose=F)

DefaultAssay(pbmc) <- "AB"
pbmc = pbmc %>%
   NormalizeData(normalization.method="CLR", margin=2, verbose=F) %>%
   FindVariableFeatures(verbose=F) %>%
   ScaleData() %>%
   RunPCA(reduction.name='ab.pca', reduction.key="abPC_", npcs=25, approx=F, verbose=F)

pbmc = pbmc %>%
   FindNeighbors(reduction='ab.pca', dims=1:20, graph.name=c('ab.nn', 'ab.snn'), do.plot=T, verbose=F) %>%
   FindClusters(graph.name="ab.snn", algorithm=3, resolution=0.3, verbose=F) %>%
   RunUMAP(reduction='ab.pca', assay='AB', dims=1:20, reduction.name="AB.umap", verbose=F, reduction.key='umap') %>%
   FindNeighbors(reduction='rna.pca', dims=1:30, graph.name=c('rna.nn', 'rna.snn'), do.plot=T, verbose=F) %>%
   FindClusters(graph.name="rna.snn", algorithm=3, resolution=0.3, verbose=F) %>%
   RunUMAP(reduction='rna.pca', assay='RNA', dims=1:30, reduction.name="RNA.umap", verbose=F, reduction.key='umap') %>%
   FindMultiModalNeighbors(reduction.list=list("rna.pca", "ab.pca"), dims.list=list(1:35, 1:20),
                            knn.graph.name="wknn", snn.graph.name="wsnn", weighted.nn.name='w.nn', verbose=F) %>%
   FindClusters(graph.name="wsnn", algorithm=3, resolution=0.07, verbose=F) %>%
   RunUMAP(reduction.name="wUMAP", verbose=F, nn.name="w.nn", reduction.key='umap')

pbmc@meta.data$cluster = pbmc@meta.data$wsnn_res.0.07
DimPlot(pbmc, reduction = "wUMAP", group.by='cluster')

pbmc <- pbmc %>% FindSubCluster(cluster='2', subcluster.name='subclust2', resolution=0.35, algorithm=3, graph.name="wknn") %>%
   FindSubCluster(cluster='0', subcluster.name='subclust0', resolution=0.35, algorithm=3, graph.name="wknn")
# DimPlot(pbmc, reduction = "wUMAP", group.by='subclust0')
pbmc@meta.data$subcluster = as.character(pbmc@meta.data$wsnn_res.0.07)
pbmc@meta.data$subcluster[pbmc@meta.data$subcluster=='2'] = pbmc@meta.data$subclust2[pbmc@meta.data$subcluster=='2']
pbmc@meta.data$subcluster[pbmc@meta.data$subcluster=='0'] = pbmc@meta.data$subclust0[pbmc@meta.data$subcluster=='0']
pbmc@meta.data$subcluster = factor(pbmc@meta.data$subcluster)
pbmc@meta.data$cluster_orig = pbmc@meta.data$cluster
# pbmc@meta.data$cluster = pbmc@meta.data$subcluster
# Idents(pbmc) = pbmc@meta.data$cluster
# svg('sc_pbmc_umap_subcluster.svg')
# DimPlot(pbmc, reduction = "wUMAP", group.by='subclust2')
# dev.off()

DimPlot(subset(pbmc, cluster_orig==2), reduction="wUMAP", group.by='subcluster')
```

## Reference mapping
### To published skin rash data
[RashX article](https://doi.org/10.1126/sciimmunol.abl9165)

```{r, include=T, eval=T, warning=F, message=F}
ref.skin <- readRDS("reference/SI_Bangert.rds")
ref.skin = Seurat::UpdateSeuratObject(ref.skin)
# DimPlot(ref.skin, reduction="umap", group.by="Ident2", label=T, label.size=3, repel=T, raster=F) + NoLegend()
genes = row.names(LayerData(pbmc, assay="RNA"))
ABs = row.names(LayerData(pbmc, assay="AB"))
ref.skin.genes = row.names(LayerData(ref.skin, assay="RNA"))
ref.skin.ABs = row.names(LayerData(ref.skin, assay="ADT"))
```

```{r, include=T, eval=T, warning=F, message=F}
pbmc = SCTransform(pbmc, verbose=F)
ref.skin = SCTransform(ref.skin, verbose=F)
anchors <- FindTransferAnchors(reference=ref.skin, query=pbmc, normalization.method="SCT",
                               reference.reduction="spca", dims=1:50, verbose=T)
pbmc <- MapQuery(anchorset=anchors, query=pbmc, reference=ref.skin, new.reduction.name='ref.skin.umap',
               refdata=list(skin="Ident2"), reference.reduction="spca", reduction.model="umap", verbose=F)
DimPlot(pbmc, reduction="wUMAP", group.by="predicted.skin", label=T, label.size=3, repel=T, shuffle=T, alpha=0.5)
rm(ref.skin); gc()

### Cell type composition based on reference mapping
ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=predicted.skin)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip()
```

### scType classification
```{r, fig.width=11, fig.height=8, warning=F, message=F}
# get cell-type-specific gene sets from our in-built database (DB)
db_ = "reference/ScTypeDB_full.xlsx"
gs_list = gene_sets_prepare(db_, "Immune system") # e.g. Immune system, Pancreas, Liver, Eye, Kidney, Brain, Lung, Adrenal, Heart, Intestine, Muscle, Placenta, Spleen, Stomach, Thymus

es.max = sctype_score(scRNAseqData=LayerData(pbmc, assay="RNA", layer="scale.data"), scaled=T, gs=gs_list$gs_positive, gs2=gs_list$gs_negative)
pbmc@meta.data$scType = factor(row.names(es.max)[apply(es.max, 2, which.max)], levels=row.names(es.max), ordered=T)
pbmc@meta.data$scCat = factor(pbmc@meta.data$scType, levels=row.names(es.max), labels=c('B', 'B', 'B', 'B', 'B', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 'NK', 'Eso', 'Neutro', 'Baso', 'Mast', 'Mono', 'Mono', 'Mono', 'Macro', 'other', 'other', 'other', 'other', 'other', 'other', 'other', 'other', 'other', 'other'), ordered=T)
normscale = function(x){(x-min(x))/(max(x)-min(x))}
pbmc@meta.data$scType.score = normscale(apply(es.max, 2, max))
pbmc@meta.data$scType.quality = normscale(apply(es.max, 2, function(x){max(x)/abs(sum(x)-max(x))}))

ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=scCat)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip()
```

### To reference PBMC dataset
[Seurat v4 reference for multimodal PBMC](https://satijalab.org/seurat/articles/multimodal_reference_mapping.html)

[original article](https://doi.org/10.1016/j.cell.2021.04.048)

```{r, include=T, eval=T, warning=F, message=F}
ref <- LoadH5Seurat("reference/pbmc_multimodal.h5seurat", verbose=F)
ref@meta.data$celltype = ref@meta.data$celltype.l1
ref@meta.data$celltype[ref@meta.data$celltype == 'NK'] = ref@meta.data$celltype.l2[ref@meta.data$celltype == 'NK']
ref@meta.data$celltype[ref@meta.data$celltype.l2 == 'ILC'] = 'ILC'

pbmc.genes = row.names(LayerData(pbmc, assay = "RNA", layer = "counts"))
pbmc.ABs = row.names(LayerData(pbmc, assay = "AB", layer = "counts"))
ref.genes = row.names(LayerData(ref, assay = "SCT", layer = "data"))
ref.ABs = row.names(LayerData(ref, assay = "ADT", layer = "counts"))
```


#### Surface markers missing in the reference data
```{r}
pbmc.ABs[!pbmc.ABs %in% ref.ABs]
```


#### Transcripts missing in the reference data
```{r}
pbmc.genes[!pbmc.genes %in% ref.genes]
```

```{r, include=T, eval=T, warning=F, message=F}
# pbmc = SCTransform(pbmc, verbose=F, assay='RNA')
ref.anchors <- FindTransferAnchors(reference=ref, query=pbmc, normalization.method="SCT",
                               reference.reduction="spca", dims=1:50, verbose=F)
pbmc <- MapQuery(anchorset=ref.anchors, query=pbmc, reference=ref, new.reduction.name='ref.pbmc.umap',
               refdata=list(pbmc.l1="celltype.l1", pbmc.l2="celltype.l2", pbmc.l3="celltype.l3", pbmc="celltype"),
               reference.reduction="spca", reduction.model="wnn.umap", verbose=F)
rm(ref); nopprint = gc(verbose=F)
DefaultAssay(pbmc) <- 'RNA'
pbmc[['prediction.score.pbmc.l1']] <- NULL
pbmc[['prediction.score.pbmc.l2']] <- NULL
pbmc[['prediction.score.pbmc.l3']] <- NULL
pbmc[['predicted_ADT']] <- NULL
```


#### Cell type composition based on reference mapping
```{r, fig.width=13, fig.height=4, warning=F}
ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=predicted.pbmc)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip()
ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=predicted.pbmc.l1)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip()
ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=predicted.pbmc.l2)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip() + scale_fill_manual(values=as.vector(t(cbind(viridis(15), rev(plasma(15))))))
```


## Seurat Clusters
For further analysis, RNA transcripts and surface markers were normalized, and reduced by PCA individually.
Afterwards, clusters are computed based on multimodal neighbors and represented by UMAP reduction.
Almost all features were found to be highly variable.
The tables show strongest regulated gene and surface marker per cluster (both positive and negative)


### UMAP
```{r, fig.width=7, fig.height=6, warning=F}
DimPlot(pbmc, reduction="wUMAP", group.by="cluster", label=T, label.size=3, repel=T, shuffle=T, alpha=0.5)
```

### Sample composition
Based on multi modal model above.
```{r, fig.width=9, fig.height=3.5}
ggplot(pbmc@meta.data, aes(x=Sample_Name, fill=cluster)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller=label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip()
```

### Annotation based on References
Reference mapping of cell types (left) and subtypes (rights) within combined (AB+RNA) Weighted Nearest Neighbor UMAP coordinates in PBMC dataset.
```{r, fig.width=11, fig.height=5, warning=F}
p1 = DimPlot(pbmc, reduction="wUMAP", group.by="predicted.pbmc", label=T, label.size=3, repel=T)
p2 = DimPlot(pbmc, reduction="wUMAP", group.by="predicted.pbmc.l2", label=T, label.size=3 ,repel=T) + NoLegend()
p1 + p2

DimPlot(pbmc, reduction="wUMAP", group.by="predicted.pbmc.l3", label=T, label.size=3, repel=T)
```

```{r, fig.width=9, fig.height=5}
pbmc@meta.data$annotation = factor(pbmc@meta.data$cluster, labels=c('C1 (CD4+ T)', 'C2 (CD8+ T)', 'C3 (NK)', 'C4 (mixed)', 'C5 (B)', 'C6 (MAIT & gdT)', 'C7 (CD14+ Mono & cDC2)', 'C8 (pDC)', 'C9 (CD16+ Mono)'))

ggplot(pbmc@meta.data, aes(x=annotation, fill=predicted.pbmc)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()

ggplot(pbmc@meta.data, aes(x=annotation, fill=predicted.skin)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()

ggplot(pbmc@meta.data, aes(x=annotation, fill=scCat)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()

# tab = with(subset(pbmc@meta.data, cluster_orig==0), table(subclust0, predicted.pbmc.l2))
# round(tab/rowSums(tab)*100, 2)

pbmc@meta.data$subannotation = factor(pbmc@meta.data$subcluster, labels=c('C1_1 (CD4+ TEM MAIT)', 'C1_2 (CD4+ TCM)', 'C1_3 (CD4+ CD8naive Treg)', 'C1_4 (CD4+ CD8 MAIT)', 'C1_5 (CD4+ gdT)', 'C1_6 (CD4+ mix)', 'C2 (CD8+ T)', 'C3 (NK1)', 'C3 (NK2)', 'C3 (NK3)', 'C3 (NK4)', 'C4 (mixed)', 'C5 (B)', 'C6 (MAIT & gdT)', 'C7 (CD14+ Mono & cDC2)', 'C8 (pDC)', 'C9 (CD16+ Mono)'))

ggplot(subset(pbmc@meta.data, cluster==2), aes(x=subannotation, fill=predicted.pbmc.l3)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()

ggplot(subset(pbmc@meta.data, cluster==2), aes(x=subannotation, fill=predicted.skin)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()

ggplot(subset(pbmc@meta.data, cluster==2), aes(x=subannotation, fill=scCat)) + geom_bar(position='fill') + 
   theme_classic() + coord_flip()


```


## Interactive visualization
```{r, fig.width=11, fig.height=7}
d = LayerData(pbmc, assay = "AB", layer = "data")
pbmc@meta.data$manual_other = d['CD4-1',] > 1 | d['CD3-1',] > 1.75 | d['CD8',] > 1 | d['CD11c',] > 2 | d['CD19',] > 1
pbmc@meta.data$manual_NK = factor(ifelse(!pbmc@meta.data$manual_other & (d['CD56-1',] > 0.5 & d['CD16',] > 0.5), 'NK', 'other'))
pbmc@meta.data$manual_ILC = factor(ifelse(!pbmc@meta.data$manual_other & (d['CD127',] > 1 & d['CD161',] > 1), 'ILC', 'other'))
pbmc@meta.data$refNK = factor(ifelse(pbmc@meta.data$predicted.pbmc.l1=='NK', 'NK', ifelse(pbmc@meta.data$predicted.pbmc.l2=='ILC', 'ILC', NA)))
pbmc@meta.data$refNKbright = factor(ifelse(pbmc@meta.data$predicted.pbmc.l2=='NK', 'NK', ifelse(pbmc@meta.data$predicted.pbmc.l2=='NK_CD56bright', 'NK bright', ifelse(pbmc@meta.data$predicted.pbmc.l2=='ILC', 'ILC', NA))))

pbmc@meta.data$Tcell = factor(ifelse(d['CD3-1',] > 1.75 & d['CD14',] < 0.5 & d['CD11c',] < 2 & d['CD19',] < 0.5 & d['IgD',] < 1 & d['IgM',] < 1, 'T cell', 'other'))
pbmc@meta.data$CD4.gated = factor(ifelse(d['CD4-1',] > 1 & d['CD3-1',] > 1.75 & d['CD8',] < 1 & d['CD11c',] < 2, 'CD4 gated', 'other'))
pbmc@meta.data$Tfh = factor(ifelse(pbmc@meta.data$CD4.gated=='CD4 gated' & d['CD185',] > 0.25 & d['CD278',] > 0.1 & d['CD279',] > 0.5, 'Tfh', 'other'))
pbmc@meta.data$CD4 = factor(ifelse(pbmc@meta.data$cluster_orig==0, 'CD4 clustered', 'other'))

Idents(pbmc) = pbmc@meta.data$CD4.gated

pbmc.cd4 = subset(pbmc, CD4.gated=='CD4 gated')
pbmc.cd4 = pbmc.cd4 %>% FindVariableFeatures(verbose=F, assay='RNA') %>% ScaleData(verbose=F, assay='RNA') %>% RunPCA(verbose=F, assay='RNA', reduction.name='RNA.pca') %>% FindVariableFeatures(verbose=F, assay='AB') %>% ScaleData(verbose=F, assay='AB') %>% RunPCA(verbose=F, assay='AB', reduction.name='AB.pca') %>% FindMultiModalNeighbors(reduction.list=list('RNA.pca', 'AB.pca'), dims.list=list(1:30, 1:(nrow(pbmc.cd4[['AB']])-1)), verbose=F, weighted.nn.name='wnn') %>% RunUMAP(nn.name='wnn', reduction.name='sub.wumap', reduction.key='wUMAP_', verbose=F) %>% FindClusters(graph.name='wsnn', algorithm=3, resolution=0.1, verbose=F, cluster.name='CD4.gated.subcluster')
pbmc.cd4@meta.data$CD4.gated.subcluster = factor(pbmc.cd4@meta.data$CD4.gated.subcluster)
pbmc.cd4@meta.data$CD4.gated.subannotation = factor(paste('blood', pbmc.cd4@meta.data$CD4.gated.subcluster))

pbmc@meta.data$CD4.gated.subcluster = 'not CD4'
pbmc@meta.data[rownames(pbmc.cd4@meta.data), ]$CD4.gated.subcluster = pbmc.cd4@meta.data$CD4.gated.subcluster
pbmc@meta.data$CD4.gated.subcluster = factor(pbmc@meta.data$CD4.gated.subcluster)

pbmc@meta.data$CD4.gated.subannotation = 'not CD4'
pbmc@meta.data[rownames(pbmc.cd4@meta.data), ]$CD4.gated.subannotation = as.character(pbmc.cd4@meta.data$CD4.gated.subannotation)
pbmc@meta.data$CD4.gated.subannotation = factor(pbmc@meta.data$CD4.gated.subannotation)

saveRDS(pbmc.cd4, 'data/pbmc.cd4.gated.RDS')


pbmc.t = subset(pbmc, Tcell=='T cell') %>%
   FindVariableFeatures(verbose=F, assay='RNA') %>%
   ScaleData(verbose=F, assay='RNA') %>%
   RunPCA(verbose=F, assay='RNA', reduction.name='RNA.pca') %>%
   FindVariableFeatures(verbose=F, assay='AB') %>%
   ScaleData(verbose=F, assay='AB') %>%
   RunPCA(verbose=F, assay='AB', reduction.name='AB.pca') %>%
   FindMultiModalNeighbors(reduction.list=list('RNA.pca', 'AB.pca'), dims.list=list(1:30, 1:(nrow(pbmc[['AB']])-1)), verbose=F, weighted.nn.name='wnn') %>%
   RunUMAP(nn.name='wnn', reduction.name='sub.wumap', reduction.key='wUMAP_', verbose=F) %>%
   FindClusters(graph.name='wsnn', algorithm=3, resolution=0.1, verbose=F, cluster.name='Tcell.subcluster')

# pbmc.t@meta.data$CD4.gated.subcluster = factor(pbmc.t@meta.data$CD4.gated.subcluster)
# tab = with(subset(pbmc.t@meta.data), table(Tcell.subcluster, predicted.pbmc.l2))
# perc = tab/rowSums(tab, na.rm=T)
# sel = apply(perc, 2, max, na.rm=T)>.01
# cbind(tab[,sel], rowSums(tab)) %>% kbl(col.names=c('# cells', colnames(tab)[sel], 'Σ')) %>% kable_classic(full_width=F)
# round(perc[,sel]*100, 1) %>% kbl(col.names=c('%', colnames(tab)[sel])) %>% kable_classic(full_width=F)
pbmc.t@meta.data$Tcell.subannotation = factor(pbmc.t@meta.data$Tcell.subcluster, labels=paste('blood', c('T_1 (mix)', 'T_2 (CD4 TCM/TEM)', 'T_3 (CD8 NK)', 'T_4 (CD4 TCM Treg)', 'T_5 (gdT ILC MAIT)', 'T_6 (mix)', 'T_7 (B)', 'T_8 (CD4 CTL CD8 TEM Treg)', 'T_9 (mix)', 'T_10 (DC CD14 Mono)', 'T_11 (mixed)')))
DimPlot(pbmc.t, reduction="sub.wumap", group.by='Tcell.subannotation', label=F, label.size=3, repel=T, shuffle=T, alpha=0.5)
# pbmc.t@meta.data$Tcell.subannotation = factor(paste('skin', pbmc.t@meta.data$Tcell.subcluster))

pbmc@meta.data$Tcell.subcluster = 'no T cell'
pbmc@meta.data[rownames(pbmc.t@meta.data), ]$Tcell.subcluster = pbmc.t@meta.data$Tcell.subcluster
pbmc@meta.data$Tcell.subcluster = factor(pbmc@meta.data$Tcell.subcluster)

pbmc@meta.data$Tcell.subannotation = 'no T cell'
pbmc@meta.data[rownames(pbmc.t@meta.data), ]$Tcell.subannotation = as.character(pbmc.t@meta.data$Tcell.subannotation)
pbmc@meta.data$Tcell.subannotation = factor(pbmc@meta.data$Tcell.subannotation)
                                   
saveRDS(pbmc.t, 'data/pbmc.tcells.RDS')

saveRDS(pbmc, 'data/pbmc.RDS')
```

The single cell analysis result is available online here:
[atlas.thlkn.de:3838/shiny-umap/](http://atlas.thlkn.de:3838/shiny-umap/)


