---
title: "PBMC statistical analysis"
always_allow_html: yes
date: "Compiled `r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    theme: united
    code_folding: hide
    collapse: yes
    css: max.css
---

[BACK TO INDEX](./index.html)


```{r, message=F, warning=F}
set.seed(123)
knitr::opts_knit$set(progress=F, verbose=F, warning=F, message=F)

subchunkify <- function(g, fig_height=7, fig_width=5, warning=F) {
  if(!isTRUE(getOption('knitr.in.progress'))){ print(g); return(); }
  g_deparsed <- paste0(deparse(function() {g}), collapse='')
  if(!exists("chunky_id")){ chunky_id <<- 0}
  chunky_id <<- chunky_id+1
  sub_chunk <- paste0("\n\n``", "`{r sub_chunky_", chunky_id, ", fig.height=", fig_height, ", fig.width=", fig_width, ", echo=F, warning=", warning, "}",
  "\n(", g_deparsed, ")()\n\n``", "`\n")
  cat(knitr::knit(text = knitr::knit_expand(text=sub_chunk), quiet=T))
}

library(Seurat)
library(tidyverse)
library(DT)
library(data.table)
library(ggplot2)
# remotes::install_github("mojaveazure/seurat-disk")
# library(SeuratDisk)
library(ggpubr)
library(rstatix)
library(celda)
library(viridis)
library(ComplexHeatmap)
library(circlize)
library(DESeq2)
library(ggrepel)
library(kableExtra)
# library(RColorBrewer)

library(future)
plan("multicore", workers=18)

meta = read.csv('meta2.csv', sep='\t', header=T, stringsAsFactors=T)
meta$PVvsHC = factor(ifelse(meta$Health %in% c('PV', 'HC'), as.character(meta$Health), NA), levels=c('PV', 'HC'))
meta$BPvsHC = factor(ifelse(meta$Health %in% c('BP', 'HC'), as.character(meta$Health), NA), levels=c('BP', 'HC'))
meta$PSORIvsHC = factor(ifelse(meta$Health %in% c('PSORI', 'HC'), as.character(meta$Health), NA), levels=c('PSORI', 'HC'))

biramp = function(n) {as.vector(t(cbind(viridis(ceiling(n/2)), colorRampPalette(c('#eedd00', '#ff5533', '#441199'), interpolate="spline", space="Lab")(ceiling(n/2)))))}

so = readRDS('data/pbmc.RDS')
meta = subset(meta, Tissue=='pbmc')
```


## UMAP from previous clustering
```{r, fig.width=7, fig.height=6}
p = DimPlot(so, reduction="wUMAP", group.by="annotation", label=F, label.size=3, repel=T, shuffle=T, alpha=0.5)
p; ggsave('sc_blood_umap_cluster.svg', plot=p, width=7, height=5)

p = DimPlot(so, reduction="wUMAP", group.by="subannotation", label=F, label.size=3, repel=T, shuffle=T, alpha=0.5)
p; ggsave('sc_blood_umap_subcluster.svg', plot=p, width=7, height=5)
```


### PBMC Reference celltypes in our UMAP
Reference mapping of cell types (left) and subtypes (rights) within combined (AB+RNA) Weighted Nearest Neighbor UMAP coordinates in PBMC dataset.
```{r, fig.width=11, fig.height=6, warning=F}
p = DimPlot(so, reduction="wUMAP", group.by="predicted.pbmc", label=F, label.size=3, repel=T, shuffle=T, alpha=0.5)
p; ggsave('sc_blood_umap_ref_pbmc.svg', plot=p, width=6, height=5)

p1 = DimPlot(so, reduction="wUMAP", group.by="predicted.pbmc.l1", label=T, label.size=3, repel=T)
p2 = DimPlot(so, reduction="wUMAP", group.by="predicted.pbmc.l2", label=T, label.size=3 ,repel=T) + NoLegend()
p1 + p2

DimPlot(so, reduction="wUMAP", group.by="predicted.pbmc.l3", label=T, label.size=3, repel=T)

tab = table(so@meta.data$refNK, so@meta.data$Health)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```

#### Cell type composition based on reference mapping
```{r, fig.width=13, fig.height=4, warning=F}
p = ggplot(so@meta.data, aes(x=Sample_Name, fill=predicted.pbmc)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip() + scale_fill_manual(values=biramp(10)) + ylab('cell composition [%]') + xlab('Samples') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_ref_pbmc.svg', plot=p, width=11, height=4)

p = ggplot(so@meta.data, aes(x=Sample_Name, fill=predicted.pbmc.l2)) + geom_bar(position='fill') + 
   facet_grid(rows=vars(Health), labeller = label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip() + scale_fill_manual(values=biramp(28)) + ylab('cell composition [%]') + xlab('Samples') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_ref_pbmc_l2.svg', plot=p, width=11, height=4)
```

### scType classification
```{r, fig.width=11, fig.height=8, warning=F, message=F}
DimPlot(so, reduction="wUMAP", group.by="scType", label=T, label.size=3, repel=T, shuffle=T)

so.umap = cbind(so@reductions$wUMAP@cell.embeddings, so@meta.data)
names(so.umap)[1:2] = c('UMAP1', 'UMAP2')
ggplot(so.umap) + geom_point(aes(UMAP1, UMAP2, col=scCat, alpha=scType.score), stroke=0, shape=16, size=1) + theme_classic() + guides(color=guide_legend(override.aes=list(alpha=1, size=3)), alpha=guide_legend(override.aes=list(size=4, shape=15))) + scale_colour_manual(values = rainbow(10))

ggplot(so.umap) + geom_point(aes(UMAP1, UMAP2, col=scType, alpha=scType.quality), stroke=0, shape=16, size=1) + theme_classic() + guides(color=guide_legend(override.aes=list(alpha=1, size=3)), alpha=guide_legend(override.aes=list(size=4, shape=15))) + scale_colour_manual(values = rainbow(33))

tab = table(so@meta.data$Health, so@meta.data$scCat)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```

#### Cells per Cluster per Celltype
```{r, fig.width=9, fig.height=3}
tab = table(so@meta.data$scCat, so@meta.data$Health)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```

#### Cross-validating cell type annotations
```{r, fig.width=9, fig.height=3}
tab = table(so@meta.data$scCat, so@meta.data$predicted.pbmc.l1)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```


## Clusters
Based on multi modal model above.
```{r, fig.width=9, fig.height=3.5}
p = ggplot(so@meta.data, aes(x=Sample_Name, fill=annotation)) + geom_bar(position='fill') + 
   scale_fill_manual(values=biramp(10)) +
   facet_grid(rows=vars(Health), labeller=label_wrap_gen(width=16, multi_line=F), scales="free", space="free") +
   theme_classic() + coord_flip() + ylab('cell composition [%]') + xlab('Samples') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_cluster.svg', plot=p, width=11, height=4)
```

### Cells per Cluster per Condition
```{r, fig.width=9, fig.height=2}
tab = table(so@meta.data$cluster, so@meta.data$Health)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```

### Cells per Cluster per Reference Celltype
```{r, fig.width=9, fig.height=3.5}
p = ggplot(so@meta.data, aes(x=annotation, fill=predicted.pbmc)) + geom_bar(position='fill') + 
   scale_fill_manual(values=biramp(10)) +
   theme_classic() + coord_flip() + ylab('cell composition [%]') + xlab('Samples') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_cluster_ref_pbmc.svg', plot=p, width=11, height=4)

tab = table(so@meta.data$annotation, so@meta.data$predicted.pbmc)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)

p = ggplot(so@meta.data, aes(x=annotation, fill=predicted.pbmc.l2)) + geom_bar(position='fill') + 
   scale_fill_manual(values=biramp(24)) +
   theme_classic() + coord_flip() + ylab('cell composition [%]') + xlab('Clusters') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_cluster_ref_pbmc_l2.svg', plot=p, width=11, height=4)

tab = table(so@meta.data$annotation, so@meta.data$predicted.pbmc.l2)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)

p = ggplot(so@meta.data, aes(x=annotation, fill=predicted.skin)) + geom_bar(position='fill') + 
   scale_fill_manual(values=biramp(24)) +
   theme_classic() + coord_flip() + ylab('cell composition [%]') + xlab('Clusters') + scale_y_continuous(breaks=(0:4)/4, labels=(0:4)*25, expand=c(0.002, 0.002)) + theme(legend.title=element_text(size=11), legend.key.height=unit(4, 'mm'), legend.key.width=unit(8, 'mm'))
p; ggsave('sc_blood_composition_cluster_ref_skin.svg', plot=p, width=11, height=4)

tab = table(so@meta.data$annotation, so@meta.data$predicted.skin)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```

```{r, fig.width=9, fig.height=3}
tab = table(so@meta.data$predicted.pbmc, so@meta.data$annotation)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)

tab = table(so@meta.data$predicted.pbmc.l2, so@meta.data$annotation)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)

tab = table(so@meta.data$predicted.skin, so@meta.data$annotation)
tabsum = tab %>% addmargins() 
tabsum %>% kbl(col.names=c('#', colnames(tabsum))) %>% kable_classic(full_width=F)
round(tab/rowSums(tab)*100, 1) %>% kbl(col.names=c('%', colnames(tab))) %>% kable_classic(full_width=F)
```


### Top markers per cluster
To characterize the clusters further, top up/down regulated features compared to the rest of the clusters are identified.


#### Transcript markers
```{r, fig.width=11, fig.height=5, warning=F, message=F}
Idents(so) = so@meta.data$annotation
so.rna.markers = FindAllMarkers(so, min.pct=0.25, verbose=F, assay='RNA')
so.rna.markers.top = so.rna.markers %>% group_by(cluster) %>% top_n(2, avg_log2FC)
so.rna.markers.top = so.rna.markers.top[, c('gene', 'cluster', 'avg_log2FC', 'p_val_adj')]
so.rna.markers.top$avg_log2FC = signif(so.rna.markers.top$avg_log2FC, digits=2)
so.rna.markers.top$p_val_adj = signif(so.rna.markers.top$p_val_adj, digits=2)

# p = DotPlot(so, features=unique(so.rna.markers.top$gene), scale.by='size', scale=T) + scico::scale_color_scico(palette = "vik") + RotatedAxis()
# p
# ggsave('sc_blood_dotplot_cluster.svg', plot=p, width=9, height=4)

DoHeatmap(so, features=unique(so.rna.markers.top$gene), size=3, slot='counts', disp.max=10) + scale_fill_viridis()
```

```{r, fig.width=11, fig.height=5, warning=F, message=F}
datatable(so.rna.markers.top, filter='top', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
```

##### Subcluster Profiles
```{r, fig.width=11, fig.height=5, warning=F, message=F}
Idents(so) = so@meta.data$subcluster
so.rna.markers = FindAllMarkers(so, min.pct=0.25, verbose=F, assay='RNA')
so.rna.markers.top = so.rna.markers %>% group_by(cluster) %>% top_n(2, avg_log2FC)
so.rna.markers.top = so.rna.markers.top[, c('gene', 'cluster', 'avg_log2FC', 'p_val_adj')]
so.rna.markers.top$avg_log2FC = signif(so.rna.markers.top$avg_log2FC, digits=2)
so.rna.markers.top$p_val_adj = signif(so.rna.markers.top$p_val_adj, digits=2)

# p = DotPlot(so, features=unique(so.rna.markers.top$gene), scale.by='size', scale=T) + scico::scale_color_scico(palette = "vik") + RotatedAxis()
# p
# ggsave('sc_blood_dotplot_subcluster.svg', plot=p, width=11, height=5)

DoHeatmap(so, features=unique(so.rna.markers.top$gene), size=3, slot='counts', disp.max=10) + scale_fill_viridis()

```


Specific expression profile compared also to all other clusters:
```{r, fig.width=7, fig.height=4, warning=F, message=F}
Idents(so) = so$subcluster
so.rna.markers = FindAllMarkers(so, min.pct=0.25, verbose=F, assay='RNA')
so.rna.markers.top = so.rna.markers %>% group_by(cluster) %>% filter(grepl('2_', cluster)) %>% top_n(2, avg_log2FC)
DotPlot(subset(so, cluster==2), features=unique(so.rna.markers.top$gene), scale.by='size', scale=T) + scico::scale_color_scico(palette = "vik") + RotatedAxis()
```

Just compared within cluster 3
```{r, fig.width=7, fig.height=3.5, warning=F, message=F}
so.rna.markers = FindAllMarkers(subset(so, cluster==2), min.pct=0.25, verbose=F, assay='RNA')
so.rna.markers.top = so.rna.markers %>% group_by(cluster) %>% filter(grepl('2_', cluster)) %>% top_n(2, avg_log2FC)
# DotPlot(subset(so, cluster==2), features=unique(so.rna.markers.top$gene), scale.by='size', scale=T) + scico::scale_color_scico(palette = "vik") + RotatedAxis()
```

##### PBMC reference annotation
```{r, fig.width=11, fig.height=5, warning=F, message=F}
Idents(so) = so@meta.data$predicted.pbmc.l1
so.rna.markers = FindAllMarkers(so, min.pct=0.25, verbose=F, assay='RNA')
so.rna.markers.top = so.rna.markers %>% group_by(cluster) %>% top_n(2, avg_log2FC)
so.rna.markers.top = so.rna.markers.top[, c('gene', 'cluster', 'avg_log2FC', 'p_val_adj')]
so.rna.markers.top$avg_log2FC = signif(so.rna.markers.top$avg_log2FC, digits=2)
so.rna.markers.top$p_val_adj = signif(so.rna.markers.top$p_val_adj, digits=2)

# p = DotPlot(so, features=unique(so.rna.markers.top$gene), scale.by='size', scale=T) + scico::scale_color_scico(palette = "vik") + RotatedAxis()
# p; ggsave('sc_blood_ref_pbmc_rnamarker_dotplot.svg', plot=p, width=11, height=5)

DoHeatmap(so, features=unique(so.rna.markers.top$gene), size=3, slot='counts', disp.max=10) + scale_fill_viridis()
```


#### Surface Markers
```{r, fig.width=7, fig.height=6}
so.ab.markers = FindAllMarkers(so, min.pct=0.25, verbose=F, assay='AB')
so.ab.markers.top = so.ab.markers %>% group_by(cluster) %>% top_n(10, avg_log2FC)
so.rna.markers.top = so.rna.markers.top[, c('gene', 'cluster', 'avg_log2FC', 'p_val_adj')]
so.rna.markers.top$avg_log2FC = signif(so.rna.markers.top$avg_log2FC, digits=2)
so.rna.markers.top$p_val_adj = signif(so.rna.markers.top$p_val_adj, digits=2)
datatable(so.rna.markers.top, filter='top', extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
```


#### Feature and violin plots of top surface markers
```{r, fig.width=13, fig.height=11, warning=F}
FeaturePlot(so, features=unique(so.ab.markers.top$gene), reduction='wUMAP')
# VlnPlot(so, features=unique(so.ab.markers.top$gene), pt.size=.05, assay='AB')
```

### Per Cluster: Disease vs HC
```{r, results='asis', message=F, warning=F}
testMarkers = function(so.sub, var, first, second, plot=T) {
   so.c.cells = table(so.sub@meta.data$Sample_Name)
   Idents(so.sub) = so.sub@meta.data[[var]]
   if(length(table(Idents(so.sub)))==1 | any(table(Idents(so.sub))[c(first, second)]<3)) { cat(paste(c('\nless than 3 cells per condition\n'))); return(NULL) }
   
   so.c.cells = table(so.sub@meta.data$Sample_Name)
   so.c.mark = FindMarkers(so.sub, test.use="MAST", ident.1=first, ident.2=second, assay='RNA', verbose=F, min.pct=0.1)
   so.c.mark = subset(so.c.mark, p_val_adj<0.1 & abs(avg_log2FC)>0.1)
   if(nrow(so.c.mark)<1) { cat(paste(c('\nno significant genes for', first, 'vs', second,'\n'))); return(NULL);}
   
   l2fc = so.c.mark[,"avg_log2FC", drop=F]
   padj = so.c.mark[,"p_val_adj", drop=F]
   padj = ifelse(padj<0.001,"***",ifelse(padj<0.01,"**",ifelse(padj<0.05,"*","")))

   p1 = Heatmap(l2fc, name="l2FC", column_gap=unit(0, "mm"), border=T, column_names_gp=gpar(fontsize=10),
                row_names_gp=gpar(fontsize=10), column_title_gp=gpar(fontsize=8), row_title_rot=0, 
                cluster_rows=T, cluster_columns=F, show_heatmap_legend=F,
                cell_fun=function(j, i, x, y, w, h, fill) { grid.text(paste0(signif(l2fc[i,j], 2), " ", padj[i,j]), x, y, gp=gpar(fontsize=7)) },
                col=colorRamp2(c(-2, 0, 2), c("#3344aa", "#eeeeee", "#aa4444")))

   pct = so.c.mark[,c("pct.1", "pct.2"), drop=F]
   colnames(pct) = c(first, second)

   p2 = Heatmap(as.matrix(pct), name="pct", column_names_gp=gpar(fontsize=10),
                row_names_gp=gpar(fontsize=12), column_title_gp=gpar(fontsize=8), column_gap=unit(0, "mm"),
                cluster_rows=T, cluster_columns=F, row_title_rot=0, border=T, show_heatmap_legend=F,
                cell_fun=function(j, i, x, y, w, h, fill) { grid.text(signif(pct[i, j], 2), x, y, gp=gpar(fontsize=10)) },
                col=colorRamp2(c(0, 1), c("#ffffff", "#555555")))
   
   avg_expr = AverageExpression(so.sub, return.seurat=F, assays="RNA", group.by="Sample_Name", slot="data", verbose=F)[[1]][rownames(so.c.mark), , drop=F]
   avg_expr = avg_expr/rowSums(avg_expr, na.rm=T)*100

   p3 = Heatmap(as.matrix(avg_expr), name="mean(exp)", row_km=2, row_names_gp=gpar(fontsize=10),
                cluster_rows=T, cluster_columns=F, show_heatmap_legend=F, column_names_gp=gpar(fontsize=10),
                column_title_gp=gpar(fontsize=8), column_gap=unit(0, "mm"), row_title_rot=0, border=T,
                cell_fun=function(j, i, x, y, w, h, fill) { grid.text(paste(round(avg_expr[i, j], 1), '%'), x, y, gp=gpar(fontsize=10)) },
                #top_annotation=columnAnnotation(cells=anno_barplot(unlist(list(so.c.cells)), height=unit(10, "mm"), border=F, add_numbers=T, gp=gpar(fill="#CCCCCC", col='white'))),
                col=colorRamp2(c(0, max(avg_expr)), c("#ffffff", "#44aa55")))
   
   if(plot) {subchunkify(p1+p2+p3, fig_height=nrow(avg_expr)/5+2, fig_width=8)}
   so.c.mark
}

c.count = table(so@meta.data$Sample_Name, so@meta.data$cluster)
perc = c.count / rowSums(c.count) * 100
for(cl in levels(so@meta.data$cluster)) {
   cat(paste(c('\n\n#### Cluster', cl, '\n')))
   so.tmp = so
   so.tmp@meta.data$cluster = factor(ifelse(so.tmp@meta.data$cluster==cl, cl, 'other'))
   subchunkify(DimPlot(so.tmp, reduction="wUMAP", group.by="cluster", label=T, label.size=3, repel=T), fig_height=5, fig_width=5)
   so.c = subset(so, cluster==cl)
   c.perc = data.frame(meta, percent=perc[as.character(meta$Sample_Name), as.character(cl)])
   c.perc.extra = rbind(c.perc, within(subset(c.perc, Health=='HC'), Health<-'BP'), within(subset(c.perc, Health=='HC'), Health<-'PV'), within(subset(c.perc, Health=='HC'), Health<-'PSORI'))
   subchunkify(datatable(compare_means(percent ~ Health, c.perc.extra, method='t.test')[,c('group1', 'group2', 'p.adj', 'p.format', 'method')]), fig_height=5, fig_width=5, warning=F)
   
   p1 = ggplot(c.perc, aes(x=Health, y=percent, fill=Health)) +
            geom_boxplot(position=position_dodge()) +
            stat_compare_means(aes(group = Health), label="p.signif") +
            theme_classic()
   # c.stat = c.perc %>% t_test(percent ~ Health, ref.group='HC') %>% add_xy_position(x = "Health", dodge = 0.8)
   c.perc <- c.perc %>% group_by(Health) %>% summarise(se=sd(percent, na.rm = TRUE)/sqrt(length(percent)), avg=mean(percent), Health=Health[1])
   p2 = ggplot(c.perc, aes(x=Health, y=avg, fill=Health)) + geom_bar(stat="identity", position=position_dodge()) +
      ylab('cells per sample [%]') +
            geom_errorbar(aes(ymin=avg-se, ymax=avg+se), width=.2, position=position_dodge(.9)) + theme_classic()# +
            # stat_pvalue_manual(c.stat, label = "p.adj", tip.length = 0.01, bracket.nudge.y = -2))
   subchunkify(p1+p2, fig_height=3, fig_width=9)
   
   for(cond in levels(factor(so@meta.data$Health))) {
      if(cond == 'HC') {next;}
      cat(paste(c('\n\n#####', cond, 'vs HC\n')))
      so.sub = subset(so, cluster==cl & Health %in% c(cond, 'HC'))
      d = testMarkers(so.sub, 'Health', cond, 'HC')
      if(is.null(d)) {next;}
      d = round(d[, c('avg_log2FC', 'p_val_adj')], digits=3)
      d$gene = row.names(d); d$signif = cut(d$p_val_adj, breaks=c(-Inf, 0.001, 0.01, 0.05, 0.10, Inf), labels=c("***", "**", "*", ".", "n.s."), right=F)
      print(htmltools::tagList(datatable(d, filter='top', extensions='Buttons', options=list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))))
   }
}
```
